---
title: "Using_speciesRanges"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using_speciesRanges}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 8,
  comment = "#>"
)
```

```{r setup}
library(speciesRanges)
```

# Tutorial to install and use speciesRanges

This tutorial aims to guide the user from the installation to the data visualisation of speciesRanges. We will walk the user through some examples. The first example uses a convex hull, produced  based on the species occurrence points, as the basis to calculate the species range. The second example uses range maps provided by the user as the base for the updated range. We ran all tests in a computer with 8GB of RAM, the execution times of each function may vary depending on the computer configuration.

## Installing speciesRanges

```{r, eval = FALSE}
install.packages ("devtools")
library(devtools)
devtools::install_github ("EduardoArle/speciesRanges")
library(speciesRanges)
```

## Example 1 - Convex hull as basis for the range map, using a simulated species.

1 - Preparing and visualising the input data. The function *occData* allows to standardise occurrence point data to feed the model.

```{r}
test_data <- data.frame(sps=rep("Equus acephalus",10),
             lon=c(-43.2,-58.4,-56,-44,-54.5,-57.4,-60.1,
                   -68.5,-71.3,-47.5),
             lat=c(-22.9,-34.6,-34.8,-20,-25.5,-25.2,-3,
                   -32.5,-41.1,-15.5),
             gender=rep("female",10),
             head_size=rep("headless individual"),
            occ=c(TRUE,TRUE,TRUE,FALSE,FALSE,TRUE,TRUE,
                  FALSE,TRUE,TRUE))

sps_occurrence <- occData(occ_table = test_data,
                          species = "sps",
                          longitude = "lon",
                          latitude = "lat",
                          occurrence = "occ")
```

Running time: ≈ 1s.
Output: a data.frame containing georeferenced records with column names matching the algorithm requirements.

```{r, warning = FALSE}
plotData(sps_occurrence)
```

Running time: ≈ 1s.
Output: a map showing the presence and absence points, respectively in blue and red. 

2 - Creating and visualising a convex hull based on the species occurrence points. The parameter *realm* indicates it is a terrestrial species, eliminating the areas of the convex hull that overlap the ocean.

```{r}
convex_hull <- convexHull(sps_occurrence, 
                          realm = "terrestrial")
```

Running time: ≈ 8s.
Output: a map showing the presence and absence points, respectively in blue and red.

```{r, warning = FALSE}
plotData(sps_occurrence, range = convex_hull)
```

Running time: ≈ 1s.
Output: a map showing the presence and absence points, respectively in blue and red; and the convex hull to be used as the basis for the range map, in green. 




## Example 2 - Specialist range map provided by the user as the base for the updated range, using *Euchloe ausonia* as an example.

1 - Preparing and visualising the input data.

```{r}
data("Euchloe_ausonia")
test_data <- Euchloe_ausonia

sps_occurrence <- occData(occ_table = test_data,
                          species = "Species",
                          longitude = "Longitude",
                          latitude = "Latitude",
                          occurrence = "occurrence")
```

Running time: ≈ 1s.
Output: a data.frame containing georeferenced records with column names matching the algorithm requirements.

```{r, warning = FALSE}
plotData(sps_occurrence)
```

Running time: ≈ 1s.
Output: a map showing the presence and absence points, respectively in blue and red.

2 - Loading and visualising the species range map.

```{r, warning = FALSE}
data("Euchloe_ausonia_shp")
plotData(sps_occurrence, range = Euchloe_ausonia_shp)
```

Running time: ≈ 1s.
Output: a map showing the presence and absence points, respectively in blue and red; and the range map to be used as the basis for the updated version, in green. 